---
title: "Semana3"
format: html
editor: visual
---

```{r}
load_pkgs <- function(pkgs) {
  for (p in pkgs) {
    if (!require(p, character.only = TRUE)) {
      install.packages(p)
      library(p, character.only = TRUE)
    }}}
load_pkgs(c("dplyr", "readxl", "ggplot2", "tidyverse", "tidyr", "car", "readxl", "stats", "MASS", "pscl"))
rm(load_pkgs) #Borrar función
setwd("/Users/mac/Desktop/AREAS/BIOL-CUANTITATIVE/random/Notes-Test/Semana3")
```

## Semana 3

Correlaciones

```{r}
plantas <- read.csv("Plant_height.csv")
glimpse(plantas)
```

```{r}

```

Regresiones

```{r}

```

### Pearson correlation

$ \mathrm{Cov}(X, Y) = \frac{1}{n-1} \sum\_{i=1}\^{n} (x_i - \bar{x})(y_i - \bar{y}) $

#### Vamos a crear nuestros datos

```{r}
data <- tibble(
  x=rnorm(100, 175,10), 
  y=rnorm(100, 70, 10)
)

shapiro.test(data$x)
shapiro.test(data$y)
```

La data es normal

```{r}
data <- data %>% 
  mutate(x_diff = x - mean(x), 
         y_diff = y - mean(y))
```

```{r}
(covariance = sum(data$x_diff*data$y_diff) / (nrow(data)-1))
```

```{r}
desv_x = sqrt((sum(data$x_diff^2)) / (nrow(data)-1))
desv_y = sqrt((sum(data$y_diff^2)) / (nrow(data)-1))
(correlation = covariance / (desv_x*desv_y))
cor.test(data$x, data$y, method = "pearson")
```

### Spearmnan correlation

$ \rho\_s = 1 - \frac{6 \sum_{i=1}^{n} d_i^2}{n(n^2 - 1)} $

```{r}
penguins <- read_csv("penguins.csv")
glimpse(penguins)
penguins <- na.omit(penguins)

data_spearman <- penguins %>% 
  dplyr::select(bill_length_mm, body_mass_g)

## Plot density

plot(density(data_spearman$bill_length_mm))
plot(density(data_spearman$body_mass_g))

## Normality test

shapiro.test(data_spearman$bill_length_mm)
shapiro.test(data_spearman$body_mass_g)

## Ranks differences

data_spearman = data_spearman %>% 
  mutate(d = rank(bill_length_mm) - rank(body_mass_g), d2 = d^2)

## Spearman's correlation

(rho = 1 - ((6 * sum(data_spearman$d2)) / 
             (nrow(data_spearman) * (nrow(data_spearman)^2 - 1))))

cor.test(data_spearman$bill_length_mm, data_spearman$body_mass_g,
         method = "spearman")

```
```{r}
# Linear regression -------------------------------------------------------
model = lm(bill_length_mm ~ body_mass_g, data = penguins)

summary(model)

### Formula: bill_len = 27.15 + 0.004 * body_mass ###

r2 = summary(model)$r.squared

set.seed(28052004)
# theme de Daniela para graficas :) 
theme <- theme_minimal(base_size = 11) + theme(
        panel.border = element_rect(color = "black", fill = NA, linewidth = 0.4),
        panel.grid = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5)
      )

ggplot(data = penguins, aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point(size = 1.5, alpha = 0.75, color="black") +
  geom_smooth(method = "lm",
              color = "red", linetype = "dashed", se = TRUE) +
  annotate("text", x = 6000, y = 40,
           label = bquote(R^2 == .(round(r2, 3)))) +
  labs(x = "Body mass (g)", y = "Bill length (mm)") +
  theme
```

```{r}

```

### Fitting Generalized Linear Models
Logistic formula 
$$\frac{1}{1 + e^{-\lambda}} $$

$$ \mathbb{P}(y = 1 | \lambda) \therefore \mathbb{P}=\frac{1}{1 + e^{-\lambda}} $$

With some math, we obtain the following expression: 

$$ \frac{1}{e^{\lambda}} = \frac{\mathbb{P}}{1- \mathbb{P}} = Odds$$
Odd ratio: calculates the relationship between a variable and probability of an event occurring.

$$Ln(\frac{\mathbb{P}}{1-\mathbb{P}}) = Ln(Odds) = \lambda$$
Therefore, we get and expression (logit function) a linear combination that represents our explanatory variable. 
```{r}
# Logit function
logit <- function(p) {
  log(p / (1 - p))
}

# Inverse logit (to go back from lambda to probability)
inv_logit <- function(lambda) {
  1 / (1 + exp(-lambda))
}


beta0 <- -1.0
beta1 <- 0.8
x <- -4:5

# Linear predictor
lambda <- beta0 + beta1 * x

# Probability of event
p <- inv_logit(lambda)

# Odds
odds <- p / (1 - p)

# Log-odds (should equal lambda)
log_odds <- logit(p)


lambda <- beta0 + beta1 * x
p <- 1 / (1 + exp(-lambda))
odds <- p / (1 - p)
log_odds <- log(p / (1 - p))

data.frame(x, lambda, p, odds, log_odds)
```

