---
title: "Clase-Semana13-GWR"
author: Daniela A. Puentes Herrera
format: html
editor: visual
---

## Load libraries
```{r warning=FALSE, message=FALSE}
library(terra)
library(tidyverse)
library(raster)
library(tmap)
library(GWmodel)
library(sf)
```

## Load the data
```{r warning=FALSE, message=FALSE}
MapSB = st_read("IncidenceMapClass.shp") # Incidencias de accidente ofidico
Temp = rast("Temp.tif") # Temperatura
Prep = rast("Prec.tif") # Precipitacion 
```

## Let's calculate the mean values of each climate variable per state (departamento)

```{r}
Tp_mean = raster::extract(Temp, MapSB, fun = mean, na.rm = TRUE)
Prep_mean = raster::extract(Prep, MapSB, fun = mean, na.rm = TRUE)
```

## Now, we can add the prep and temp data to MapSB
```{r}
MapSB$Prec = as.numeric(Prep_mean$Prec)
MapSB$Temp = as.numeric(Tp_mean$Temp)
```

##  Save the .shp 
```{r}
st_write(MapSB, "/Users/mac/Desktop/AREAS/BIOL-CUANTITATIVE/random/Notes-Test/MapaFClass.shp", driver = "ESRI Shapefile", append = FALSE)
```

## Load the new data 

```{r}
rm()
# Map = st_read("/Users/mac/Desktop/AREAS/BIOL-CUANTITATIVE/random/Notes-Test/MapaFClass.shp") # Incidencias de accidente ofidico

# Method 1: Using terra::vect()
map <- terra::vect("/Users/mac/Desktop/AREAS/BIOL-CUANTITATIVE/random/Notes-Test/MapaFClass.shp")


```

## Ordinary Linear Regression  

```{r}
ModLM=lm(data=as.data.frame(Map),formula=InTot~Prec+Temp)
summary(ModLM)
```
## Do we have any spatial structure? What about the residual of our previous model? 

```{r}
#Calculamos residuales y los asignamos al mapa
Map$RES=as.numeric(residuals(ModLM))
#Gráficamos el mapa 
tm_shape(st_as_sf(Map))+tm_polygons(col="RES",style="pretty",palette="-RdBu")+tm_legend(outside=TRUE,text.size=0.8)
```
Positive values are associates with overfitting

## We have a spatial structure. Now, what? 

```{r}
#Calculando las coordenadas del centroide por depto (Distancias -> Wi,j)
centr=terra::centroids(spatvector_object)
```



