---
title: "Clase-Semana13-GWR"
author: Daniela A. Puentes Herrera
format: html
editor: visual
---

## Load libraries
```{r warning=FALSE, message=FALSE}
library(terra)
library(tidyverse)
library(raster)
library(tmap)
library(GWmodel)
library(sf)
```

## Load the data
```{r warning=FALSE, message=FALSE}
MapSB = st_read("IncidenceMapClass.shp") # Incidencias de accidente ofidico
Temp = rast("Temp.tif") # Temperatura
Prep = rast("Prec.tif") # Precipitacion 
```

## Let's calculate the mean values of each climate variable per state (departamento)

```{r}
Tp_mean = raster::extract(Temp, MapSB, fun = mean, na.rm = TRUE)
Prep_mean = raster::extract(Prep, MapSB, fun = mean, na.rm = TRUE)
```

## Now, we can add the prep and temp data to MapSB
```{r}
MapSB$Prec = as.numeric(Prep_mean$Prec)
MapSB$Temp = as.numeric(Tp_mean$Temp)
```

##  Save the .shp 
```{r}
st_write(MapSB, "/Users/mac/Desktop/AREAS/BIOL-CUANTITATIVE/random/Notes-Test/MapaFClass.shp", driver = "ESRI Shapefile", append = FALSE)
```

## Load the new data 

```{r}
rm()
# Map = st_read("/Users/mac/Desktop/AREAS/BIOL-CUANTITATIVE/random/Notes-Test/MapaFClass.shp") # Incidencias de accidente ofidico

# Method 1: Using terra::vect()
map <- terra::vect("/Users/mac/Desktop/AREAS/BIOL-CUANTITATIVE/random/Notes-Test/MapaFClass.shp")
```

## Ordinary Linear Regression  

```{r}
ModLM=lm(data=as.data.frame(Map),formula=InTot~Prec+Temp)
summary(ModLM)
```
## Do we have any spatial structure? What about the residual of our previous model? 

```{r}
#Calculamos residuales y los asignamos al mapa
Map$RES=as.numeric(residuals(ModLM))
#Gráficamos el mapa 
tm_shape(st_as_sf(Map))+tm_polygons(col="RES",style="pretty",palette="-RdBu")+tm_legend(outside=TRUE,text.size=0.8)
```
Positive values are associates with overfitting

## We have a spatial structure. Now, what? 

The observations are present in the current data set. But, the distances between them should be calculating using the centroids distance. 

```{r}
#Calculando las coordenadas del centroide por depto (Distancias -> Wi,j)
centr=terra::centroids(spatvector_object)
```

The weights are a function of distance. First, which are the coordinates of the previous centroids? Second, save those coordinates in a shared dataframe. 

```{r}
#Matriz distancias
cords1= crds(centr, df = TRUE)
cords2=data.frame(x=cords1[[1]],y=cords1[[2]])
```

Now, distances between all coordiantes can be calculates using `gw.dist()`
```{r}
cords=as.matrix(cords2)
DM=gw.dist(dp.locat=cords)
dim(DM)
```

## We move on to the second step which is to define what bandwidth or bandwidth we are going to use.
```{r}
KR="gaussian" #Definimos tipo de kernel osea de ancho de banda
cords2=cbind(cords2,as.data.frame(spatvector_object))
```

```{r}
coordinates(cords2) = c("x", "y")
```

The bandwidth can be calculated using `bw.gwr()`
```{r}
bw1=bw.gwr(data=cords2,formula=InTot~Prec+Temp,dMat=DM,kernel=KR,
           adaptive=T,approach="AIC")
```

## GWR model 
```{r}
#Tenemos el ancho de banda mejor, vamos a correr GWR con todas las posibles combinaciones entre mis variables independientes
gwrmodels=gwr.model.selection(DeVar="InTot",InDeVars=c("Temp","Prec"), data=cords2,bw=bw1,approach="AIC",kernel=KR,adaptive=T)

gwrmodels
```

```{r}
#Mejor modelo con ambas variables, pero es muy similar al modelo solo con temperatura
#Modelo ambas variables
Mod1av=gwr.basic(data=cords2,formula=InTot~Temp+Prec,bw=bw1,
                 dMat=DM,kernel=KR,adaptive=T)
Mod1av
```

```{r}
#Sacamos los resultados del GWR 
RESULT=Mod1av[["SDF"]]
head(RESULT@data)
```


```{r}
spatvector_object$R2=RESULT$Local_R2

#Graficamos
tm_shape(st_as_sf(spatvector_object))+tm_polygons(col="R2",style="pretty",palette="-RdBu")+tm_legend(outside=TRUE,text.size=0.8)
```
```{r}
pvals=gwr.t.adjust(Mod1av)
pp=pvals[["SDF"]] #Sacamos el mapa de pvalues
head(pp@data) #Revisar correcciones del pvalue calculada
```

```{r}
spatvector_object$PREC=RESULT$Prec*(pp$Prec_p<0.1)
tm_shape(st_as_sf(spatvector_object))+tm_polygons(col="PREC",style="pretty",palette="-RdBu")+tm_legend(outside=TRUE,text.size=0.8)
```

```{r}
spatvector_object$TEMP=RESULT$Temp*(pp$Temp_p<0.1)
tm_shape(st_as_sf(spatvector_object))+tm_polygons(col="TEMP",style="pretty",palette="-RdBu")+ tm_legend(outside=TRUE,text.size=0.8)
```

